DATA 621 - Business Analytics and Data Mining

### 2020-12-11 Final_project/Final Project Paper - Group 2.Rmd
add another GLM, remove exceed mean feature, minor text additions

The second is `exceedMean`. This will be a factor if the views for the video
are greater than the *mean* views for all of the extracted data set. This will
also indicate the skewness of the data. Were it symmetrical, and we know it is
not from the graphs above, the mean should be near the median and this factor
should be close to equal for both 0 and 1.

@@

# Add the engineered feature
usExtract[, `:=`(balance = ifelse(rtgDisabled, 1,
                               pmin(likes, dislikes) / pmax(likes, dislikes)),
                 exceedMean = factor(ifelse(views > usMean, 1, 0),
                                     levels = c(0, 1), labels = c(0, 1)))]
```
```{r plotBalance, fig.height=4, fig.width=8}
ggplot(usExtract[, .(balance)], aes(x = balance)) +
  geom_density() +
  ggtitle("Distribution of Balance")
```

@@

```{r plotExceedMean, fig.height=4, fig.width=8}
# Melt for exceedMean
usSetF <- melt(usExtract, measure.vars = 'exceedMean', variable.factor = FALSE,
               variable.name = 'measure', value.name = 'value')

# Plot the exceedMean variable
ggplot(usSetF[measure == 'exceedMean'], aes(x = value)) +
  geom_bar() +
  ggtitle("Proportion Above and Below Mean") +
  scale_y_continuous(labels = comma)
```

As with many events which follow the Pareto rule, most of the views are 
attributable to a minority of the videos, as seen by the 0/1 disparity.

@@

```{r dummyVarsAA}
# Create dummy Variables. These will be used for the next three models
modDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
                   cmtDisabled + rtgDisabled + vidError + balance,
                   data = seenSet, fullRank = TRUE)
seenX <- predict(modDum, seenSet)
seenY <- seenSet$views
```

@@

### Generalized Linear Model: Gaussian Errors & Log Link
The generalized linear model (GLM) is an extension of the simple linear model,
but the errors can be distributed per any member of the exponential family and
the relationship between some function of the predictors---called the link---and
the mean needs to be linear, not that the mean itself must be linear in the
predictors [@Faraway2006EtL].

The model under consideration here assumes a Gaussian distribution of the
errors, but a multiplicative relationship between the mean and the predictors.
This is expressed by using a log link function. This is **not** the canonical
link function for the Gaussian GLM, but as we are using numerical methods there
is no issue.

@@

\footnotesize
```{r lm2Table}
kable(summary(lm2$finalModel)$coefficients,
      caption = "Model 3 Linear Regression Output",
      digts = 3L, format.args = list(big.mark = ','))
```
\normalsize

This model has more predictors than does the simple Gaussian (G)LM. Moreover,
the signs are different from the simple LM. The very fact that
`Nonprofits & Activism` has a positive coefficient should raise concerns. This
model is probably a poor one.

@@

```{r modTest}
# Process the hidden set
tstDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
                   cmtDisabled + rtgDisabled + vidError + balance,
                   data = hideSet, fullRank = TRUE)
hideX <- predict(tstDum, hideSet)
hideY <- hideSet$views

# Predict using the models
lm1P <- predict(lm1, newdata = hideX)
lm2P <- predict(lm2, newdata = hideX)
lm3P <- predict(lm3, newdata = hideX)

# Compare the results
compTable <- data.table(Model = c('LM', 'GLM: Gauss+Log', 'ElasticNet'),
                        RMSE = c(RMSE(lm1P, hideY), RMSE(lm2P, hideY),
                                 RMSE(lm3P, hideY)),
                        R2 = c(R2(lm1P, hideY, formula = 'traditional'),
                               R2(lm2P, hideY, formula = 'traditional'),
                               R2(lm3P, hideY, formula = 'traditional')),
                        MAE = c(MAE(lm1P, hideY), MAE(lm2P, hideY),
                                MAE(lm3P, hideY)))
kable(compTable, digits = 3L, format.args = list(big.mark = ','),
      caption = "Model Performance on Test Set")
```

>>>>

<!-- The second is `exceedMean`. This will be a factor if the views for the video -->
<!-- are greater than the *mean* views for all of the extracted data set. This will -->
<!-- also indicate the skewness of the data. Were it symmetrical, and we know it is -->
<!-- not from the graphs above, the mean should be near the median and this factor -->
<!-- should be close to equal for both 0 and 1.  -->

The second is `engagement`. This will be the ratio of comments to sum of likes and 
dislikes, which should give us some measure of comments to ratings. We will use a
similar approach as `balance` in terms of constraining the interval to \([0,1])\).

@@

# Add the engineered feature
usExtract[, `:=`(balance = ifelse(rtgDisabled, 1,
                               pmin(likes, dislikes) / pmax(likes, dislikes)),
                 engagement = ifelse(rtgDisabled, 1, 
                                     comments / sum(likes, dislikes)))]
                 # exceedMean = factor(ifelse(views > usMean, 1, 0),
                 #                     levels = c(0, 1), labels = c(0, 1)))

```
```{r plotBalance, fig.height=4, fig.width=8}
ggplot(usExtract[, .(balance)], aes(x = balance)) +
  geom_density() +
  ggtitle("Distribution of Balance")

ggplot(usExtract[, .(engagement)], aes(x = engagement)) +
  geom_density() +
  scale_x_log10() +
  ggtitle("Distribution of Engagement Log Scale")

```

@@

<!-- ```{r plotExceedMean, fig.height=4, fig.width=8} -->
<!-- # Melt for exceedMean -->
<!-- usSetF <- melt(usExtract, measure.vars = 'exceedMean', variable.factor = FALSE, -->
<!--                variable.name = 'measure', value.name = 'value') -->

<!-- # Plot the exceedMean variable -->
<!-- ggplot(usSetF[measure == 'exceedMean'], aes(x = value)) + -->
<!--   geom_bar() + -->
<!--   ggtitle("Proportion Above and Below Mean") + -->
<!--   scale_y_continuous(labels = comma) -->
<!-- ``` -->

<!-- As with many events which follow the Pareto rule, most of the views are  -->
<!-- attributable to a minority of the videos, as seen by the 0/1 disparity. -->

@@

```{r dummyVarsAA}
# Create dummy Variables. These will be used for the next three models
modDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
                   cmtDisabled + rtgDisabled + vidError + balance + engagement,
                   data = seenSet, fullRank = TRUE)
seenX <- predict(modDum, seenSet)
seenY <- seenSet$views
```

@@

### Generalized Linear Model: Gaussian and Poisson Errors with Log Link
The generalized linear model (GLM) is an extension of the simple linear model,
but the errors can be distributed per any member of the exponential family and
the relationship between some function of the predictors---called the link---and
the mean needs to be linear, not that the mean itself must be linear in the
predictors [@Faraway2006EtL].

The models under consideration here assume either a Poisson or Gaussian 
distribution of the errors, but a multiplicative relationship between the
mean and the predictors. This is expressed by using a log link function. 
This is **not** the canonical link function for the GLM distributions,
but as we are using numerical methods there is no issue.

@@

```{r lm3Train_poi}
set.seed(9450)
lm3_test<- train(x = seenX, y = seenY, family = poisson(link='log'),
            method = 'glmStepAIC', direction = 'both', trace = 0,
            trControl = trC)
```

\footnotesize
```{r poi_coeff}
kable(summary(lm3_test$finalModel)$coefficients,
      caption = "Model 3 (Poisson) Linear Regression Output",
      digts = 3L, format.args = list(big.mark = ','))
```
\normalsize

\footnotesize
```{r lm2Table}
kable(summary(lm2$finalModel)$coefficients,
      caption = "Model 3 (Gaussian) Linear Regression Output",
      digts = 3L, format.args = list(big.mark = ','))
```
\normalsize

These models have more predictors than the simple Gaussian (G)LM. Moreover,
the signs are different from the simple LM. The very fact that
`Nonprofits & Activism` has a positive coefficient should raise concerns. These
models are probably poor ones.

@@

```{r modTest}
# Process the hidden set
tstDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
                   cmtDisabled + rtgDisabled + vidError + balance + engagement,
                   data = hideSet, fullRank = TRUE)
hideX <- predict(tstDum, hideSet)
hideY <- hideSet$views

# Predict using the models
lm1P <- predict(lm1, newdata = hideX)
lm2P <- predict(lm2, newdata = hideX)
lm3P <- predict(lm3, newdata = hideX)
lm3_testP <- predict(lm3_test, newdata = hideX)
# Compare the results
compTable <- data.table(Model = c('LM', 'GLM: Gauss+Log', 'ElasticNet', 'GLM: Poisson+Log '),
                        RMSE = c(RMSE(lm1P, hideY), RMSE(lm2P, hideY),
                                 RMSE(lm3P, hideY), RMSE(lm3_testP, hideY)),
                        R2 = c(R2(lm1P, hideY, formula = 'traditional'),
                               R2(lm2P, hideY, formula = 'traditional'),
                               R2(lm3P, hideY, formula = 'traditional'),
                               R2(lm3_testP, hideY, formula = 'traditional')),
                        MAE = c(MAE(lm1P, hideY), MAE(lm2P, hideY),
                                MAE(lm3P, hideY), MAE(lm3_testP, hideY)))
kable(compTable, digits = 3L, format.args = list(big.mark = ','),
      caption = "Model Performance on Test Set")
```
